Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2015-01-27 windheat.tex
appendix

For a stellar population that forms impulsively of age $t$, the rate of mass and energy injection per unit stellar mass are given, respectively, by 
\begin{align} 
  \dot{\bar{m}}(t) &= \frac{\Delta M(t) \mu|_{M_{\rm TO}(t)}
    \left|\dot{M}_{\rm TO}(t)\right| + f_{\rm MS} \int_{m_0}^{m_{\rm
        T}(t)}
    \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar }{\bar{m}_*}\\
  \dot{\bar{e}}(t) &=\dot{e}_{\rm TO}(t)+ f_{\rm MS} \int_{m_0}^{m_{\rm T}(t)}
  \frac{\vw^2(\Mstar, t) \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar}{\bar{m}_*},
  \label{eq:edotImp}
\end{align} 
where the first and second terms in each expression corresponds to contributions from main sequence stars and post-main sequence (PMS) stars, respectively.  Here $ f_{\rm MS} < 1$ is an efficiency, which accounts for the efficiency with which main sequence winds thermalize their energy with the rest of the injected gas (we hereafter assume $f_{\rm MS} = 1$) and $\mu$ is the IMF, which we take to be of the Salpeter form $\mu(\Mstar)\sim M^{-2.35}$, truncated at $m_0 = 0.1 \Msun$ on the low mass end and at $100 \Msun$ on the high mass end, where $\bar{m}_*$ = 0.35 $\Msun$ is the corresponding mean stellar mass.  The turn-off mass $M_{\rm TO}$ at time $t$ is calculated according to the following functional fit
\begin{align}
\log(M_{\rm TO})/M_{\odot} =0.24 + 0.068 x^2-0.34 x+4.76 e^{-4.58 x},
\end{align}
where $x=\log(t/10^9 {\rm years})$, so designed to reproduce the results of \citet{MaederMeynet:1987a} (their Table 9) for massive stars while asymptoting to the fitting formula provided by \citet{CiottiOstriker:2007a} (their eq.~[9]) for intermediate and late times ($t \gsim 10^8$ years).

The mass lost in PMS winds $\Delta M(t)$ is calculating from \citet{CiottiOstriker:2007a} (their eq.~[10]) according to
\begin{align}
\Delta M=
\begin{cases}
0.945 M_{\rm TO}-0.503 & M_{\rm TO} < 9 \Msun\\
 M_{\rm TO}-1.4 \Msun &  M_{\rm TO} \ge 9 \Msun
\end{cases}
\end{align}
The MS wind mass loss rate $\dot{m}(\Mstar, t)$ is calculated based on the generalization of Reimer's law given in \citet{SchroderCuntz:2005a} (their eq.~[4]):
\begin{align}
  \dot{m}(\Mstar)=8 \times 10^{-14} \frac{L_* r_*}{\Mstar}
  \left(\frac{T_{\rm eff}}{\rm 4000 K}\right)^{3.5}
  \left(1+\frac{g_{\odot}}{4300 g_*}\right) \Msun \pyear,\
\end{align}
where  $R_*$, $L_*$, $T_{\rm eff}$ and $g_*$ are the stellar radius,
luminosity, effective temperature, and surface gravity, respectively, the latter normalized to its stellar value $g_{\odot}$.  The stellar radius and luminosity are estimated as 
\begin{align}
L_*=
\begin{cases}
L_{\odot} (\Mstar/\Msun)^{3.2} & \Mstar > \Msun \\
L_{\odot} (\Mstar/\Msun)^{2.5} & \Mstar \le \Msun
\end{cases}
\end{align}
\begin{align}
R_*=
\begin{cases}
R_{\odot} (\Mstar/\Msun)^{0.57} & \Mstar > \Msun \\
R_{\odot} (\Mstar/\Msun)^{0.8} & \Mstar \le \Msun
\end{cases}
\end{align}
based on \citealt{Kippenhahn&Weigert} (their Figs.~22.2, 22.3)

To get a handle on $\dot{e}_{\rm TO}(t)$, we use the results from
\citet{VossDiehl+:2009a} who use a population synthesis code to
simulate the mass and energy injection into the ISM from an OB
association. For the first $\sim 10$ Myr the energy injection will
be dominated by fast Wolf-Rayet star winds. The energy injection rate
per massive star ($\Mstar>8 \Msun$), $\dot{\mathcal{E}} (t)$, from
stellar winds is plotted in the top panel of their Figure 7 and is
well fit {\bf AG: A little more detail here...} by

\begin{align}
\dot{\mathcal{E}} (t)=
\begin{cases}
  1.3 \times 10^{36} {\rm ergs/s} & t<4 \times 10^6 {\rm years}\\
  1.3  \times 10^{36} {\rm ergs/s} \left(\frac{t}{4 \times  10^6}\right)^{-3.73} & t \ge 4 \times 10^6 {\rm years}.
\end{cases}
\label{eq:voss}
\end{align}

Equation~\eqref{eq:voss} is valid only for $t \lsim 10$ Myr. For $t\gsim
10$ Myr, $\dot{e}_{\rm TO}$ will be negligible as lower mass stars
shed their gas is slow dust-driven winds during their post-main
sequence evolution {\bf AG-Nick is the preceding statement accurate},
and for our purposes the exact functional form of $\dot{e}_{\rm
  TO}(t)$ is unimportant. So we adopt Equation~\eqref{eq:voss} for all times.

 $\dot{e}_{\rm TO}(t)$ is related to $\dot{\mathcal{E}}$  via

\begin{align}
\dot{e}_{\rm TO}(t)=f_{8} \dot{\mathcal{E}} / \bar{m}_*,
\label{eq:eto}
\end{align}

where $f_{8}$ is the fraction of the stellar population with $\Mstar>8
\Msun$. $f_8=2.6 \times 10^{-3}$ for our assumed Salpeter IMF.

We calculate the wind velocity for main sequence winds $v_w (\Mstar,
t)$ using...{\bf AG:Current just use the value for the sun~430 km/s
  replace with the real prescription we end up using.}

The effective wind velocity in the impulsive limit may then be written
as

\begin{align}
\bar{v}_w(t)=2 \dot{\bar{e}}(t)/\dot{\bar{m}}(t)
\label{eq:vwImp}
\end{align}
while {\bf BDM: insert correct expression}
\begin{align}
\eta = \dot{\bar{m}}(t)
\label{eq:vwImp}
\end{align}

We plot $\bar{v}_w$ versus time in Figure~\ref{fig:vwImp}.

\begin{figure}
\includegraphics[width=\columnwidth]{vwImp.pdf}
\caption{\label{fig:vwImp} The effective $\vw$ from stellar winds from
  a stellar population formed in a starburst $t$ years ago.}
\end{figure}

\subsection{Over Star Formation History}

We can then use these integrated quantities to determine the effective
$V_w$ for arbitrary star formation histories. For a stellar population
with star formation rate $S(t)$

\begin{align} 
  \dot{M}(t) &= \int_0^t S(t_1) \dot{\bar{m}}(t-t_1){\rm
      d}t_1\\
  \dot{E}(t) &= \int_0^t S(t_1) \dot{\bar{e}}(t-t_1){\rm
      d}t_1\\
  V_w^2(t) &=2 \dot{E}(t)/\dot{M}(t)
\end{align}

@@

We show $V_{\rm W}$ for different star formation histories in
Fig. \ref{NickPlot2} {\bf AG: this plot still uses the old ad hoc vw
  prescription? If so it would need to be fixed. In any case, we still
  differ in detail on the vw from stars}.  In particular, we use
Eqs. 17-20 from \citet{MosterNaab+:2013a} and the $M_{\rm BH}-M_{\rm
  halo}$ relation from \citet{BandaraCrampton+:2009a} to define $S(t)$
for particular galaxies.  It seems that ``bumpy'' SF histories
severely diminish $V_{\rm w}$ if the timescale for SF variability is a
factor of a few or more greater than the duration of high-velocity
winds (either 10 or 40 Myr).

>>>>

For a stellar population that forms impulsively, the rate of mass and energy injection per unit stellar mass at age $t$ are given, respectively, by 
\begin{align} 
  \dot{\bar{m}}(t) &= \frac{\Delta M(t) \mu|_{M_{\rm TO}(t)}
    \left|\dot{M}_{\rm TO}(t)\right| + f_{\rm MS} \int_{m_0}^{m_{\rm
        T}(t)}
    \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar }{\bar{m}_*}\\
  \dot{\bar{e}}(t) &=\dot{e}_{\rm TO}(t)+ f_{\rm MS} \int_{m_0}^{m_{\rm T}(t)}
  \frac{\vw^2(\Mstar, t) \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar}{\bar{m}_*},
  \label{eq:edotImp}
\end{align} 
where the first and second terms in each expression corresponds to contributions from main sequence (MS) stars and post-main sequence (PMS) stars, respectively.  Here $ f_{\rm MS} < 1$ is the efficiency with which main sequence winds thermalize their energy with the rest of the injected gas (we hereafter assume $f_{\rm MS} = 1$) and $\mu$ is the IMF, which we take to be of the Salpeter form $\mu(\Mstar)\sim M^{-2.35}$, truncated at $m_0 = 0.1 \Msun$ on the low mass end and at $100 \Msun$ on the high mass end, where $\bar{m}_*$ = 0.35 $\Msun$ is the corresponding mean stellar mass.

The quantity of mass lost in PMS winds $\Delta M(t)$ is estimated from the expression given by \citet{CiottiOstriker:2007a} (their eq.~[10]),
\begin{align}
\Delta M=
\begin{cases}
0.945 M_{\rm TO}-0.503 & M_{\rm TO} < 9 \Msun\\
 M_{\rm TO}-1.4 \Msun &  M_{\rm TO} \ge 9 \Msun,
\end{cases}
\end{align}
where $M_{\rm TO}$ is the turn-off mass, which at time $t$ is calculated as
\begin{align}
\log(M_{\rm TO})/M_{\odot} =0.24 + 0.068 x^2-0.34 x+4.76 e^{-4.58 x},
\end{align}
where $x=\log(t/10^9 {\rm years})$.  This functional fit is designed to reproduce the results of \citet{MaederMeynet:1987a} (their Table 9) for massive stars while asymptoting to the formula provided by \citet{CiottiOstriker:2007a} (their eq.~[9]) for intermediate and late times ($t \gsim 10^8$ years).

The MS wind mass loss rate $\dot{m}(\Mstar, t)$ is calculated based on the generalization of Reimer's law given in \citet{SchroderCuntz:2005a} (their eq.~[4]) as
\begin{align}
  \dot{m}=8 \times 10^{-14} \frac{L_* R_*}{\Mstar}
  \left(\frac{T_{\rm eff}}{\rm 4000 K}\right)^{3.5}
  \left(1+\frac{g_{\odot}}{4300 g_*}\right) \Msun \pyear,\
\end{align}
where  $R_*$, $L_*$, $T_{\rm eff}$ and $g_*$ are the stellar radius,
luminosity, effective temperature, and surface gravity, respectively, the latter normalized to its solar value $g_{\odot}$.  The stellar radius and luminosity are estimated as (\citet{Kippenhahn&Weigert90}; Figs.~22.2) 22.3)
\begin{align}
L_*=
\begin{cases}
L_{\odot} (\Mstar/\Msun)^{3.2} & \Mstar > \Msun \\
L_{\odot} (\Mstar/\Msun)^{2.5} & \Mstar \le \Msun
\end{cases}
\end{align}
\begin{align}
R_*=
\begin{cases}
R_{\odot} (\Mstar/\Msun)^{0.57} & \Mstar > \Msun \\
R_{\odot} (\Mstar/\Msun)^{0.8} & \Mstar \le \Msun
\end{cases}
\end{align}
The wind velocity of main sequence winds $v_w (\Mstar, t)$ is assumed to equal 430 km s$^{-1}$, corresponding to the escape speed of the Sun; this produces an effective wind heating velocity for main sequence winds alone of $XXX$, in agreement with the value of 150 km s$^{-1}$ found by \citet{NaimanSoares-Furtado+:2013a} based on a more sophisticated population synthesis treatment. {\bf BDM: true? correct}...{\bf AG:Current just use the value for the sun~430 km/s
  replace with the real prescription we end up using.}

The rate of energy input from PMS winds, $\dot{e}_{\rm TO}(t) = f_{8}\dot{\mathcal{E}}/\bar{m}_*$, is dominated by fast Wolf-Rayet winds and core-collapse supernovae from massive stars ($M_{\star} > 8M_{\odot}$), where $f_{8} =2.6 \times 10^{-3}$ is the fraction of the stellar mass with $M_{\star} > 8M_{\odot}$ for our assumed Salpeter IMF.  Here $\dot{\mathcal{E}} (t)$ is the energy injection rate per massive star, which we estimate as
\begin{align}
\dot{\mathcal{E}} (t)=  1.3 \times 10^{36} {\rm erg\,s^{-1}}
\begin{cases}
  1 & t<4 \times 10^6 {\rm yr}\\
  \left(\frac{t}{4 \times  10^6\,{\rm yr}}\right)^{-3.73} & t \ge 4 \times 10^6 {\rm yr},
\end{cases}
\label{eq:voss}
\end{align}
based on the results of \citet{VossDiehl+:2009a} (their Fig.~7, top panel), who use a population synthesis code to simulate the mass and energy injection into the ISM from an OB association.  Although equation~\eqref{eq:voss} is valid only for $t \lsim 10$ Myr, in practice the precise functional form of $\dot{e}_{\rm TO}(t)$ is generally unimportant for our purposes, so we adopt equation~\eqref{eq:voss} as being valid for all times.

The effective wind velocity in the limit of an impulsive star formation may then be written
as 
\begin{align}
\bar{v}_w(t)=2 \dot{\bar{e}}(t)/\dot{\bar{m}}(t)
\label{eq:vwImp}
\end{align}
while {\bf BDM: insert correct expression}
\begin{align}
\eta = \dot{\bar{m}}(t)
\label{eq:vwImp}
\end{align}
We plot $\bar{v}_w(t)$ and $\eta(t)$ in Figures~\ref{fig:vwImp} and XXX.

\begin{figure}
\includegraphics[width=\columnwidth]{vwImp.pdf}
\caption{\label{fig:vwImp} The effective $\vw$ from stellar winds from
  a stellar population of age $t$.}
\end{figure}

\subsection{Over Star Formation History}
Generalizing to an arbitrary star formation history $S(t)$, the total rate of mass and energy input can be written as
\begin{align} 
  \dot{M}(t) &= \int_0^t S(t_1) \dot{\bar{m}}(t-t_1){\rm
      d}t_1\\
  \dot{E}(t) &= \int_0^t S(t_1) \dot{\bar{e}}(t-t_1){\rm
      d}t_1,\\
\end{align}
resulting in an effective wind heating rate of
\begin{align}
  v_w^2(t) &=2 \dot{E}(t)/\dot{M}(t)
\end{align}

@@

We estimate the stellar wind heating provided by the {\it average} star formation history of galaxies of a given mass using the results of \citet{MosterNaab+:2013a} (their eqs.~[17]-[20]) and the $M_{\bullet}-M_{\rm
  halo}$ relation from \citet{BandaraCrampton+:2009a}.  Figure \ref{NickPlot2} shows the resulting $v_{\rm W}$ as a function of SMBH mass.  For comparison we also show cases in which the star formation rate is modulated sinuisoidally from its average value.   It seems that ``bumpy'' SF histories severely diminish $v_{\rm w}$ if the timescale for SF variability is a factor of a few or more greater than the duration of high-velocity winds (either 10 or 40 Myr). {\bf AG: this plot still uses the old ad hoc vw
  prescription? If so it would need to be fixed. In any case, we still
  differ in detail on the vw from stars}.
